---
title: "cleaning_isotope_data"
author: "Beth Davis"
date: "8/9/2021"
output: html_document
---

## Description and Metadata Information 

  This information has been preserved from notes left in the raw files before cleaning and will be listed with the sheets they apply to under the raw file name.
  
Un-Edited Notes from Raw:
    
  File: 2007-08_Palmyra_Gecko Isotopes
  Individual records - half indicate that the reads came back in May 2008, so I       assume those were geckos collected in 2007? The other half are reads that came     back in 2009, so these are presumed to be from 2008.
    
  File: 2009_Palmyra_Crab_Isotopes
  Individual records - coconut crabs, orange cardisoma, and both species hermit        crabs
  I'm a little confused about the last few entries labeled "coconut", there are no     coconut crabs on lost, so I'm thinking these were Cocos trees?


  File: 2009-2012_Palmyra_Plant_Isotopes
  * plant isotope data	2010-2012	pooled by islet within year

  File: 2008_Palmyra_Rat_Isotopes
  * 2009	rat isotope data	individuals
    
  File: 2009_Palmyra_Cane_Spider_Isotopes
  * Cane spider isotopes	2009	Individuals

  File: 2009_Palmyra_Day_Spider_Isotopes
  2009	small spider isotopes	pooled	I assume this means Neoscona theisi data. It     is pooled by island among a few individuals

  File: 2010_Palmyra_Cane_spider_Isotopes
  cane spider isotope data	2010	pooled per island for 2-15 individuals for each

  File: 2009_Palmyra_Insect_Amphipod_Isopod_Isotopes
  2009	insect, amphipod, and isopod isotope data	individuals	not sure on species    names for these… or whether they were one particular species. Ask Hillary
  
  File: 2007-2009_Palmyra_Gecko_Sizes_and_Isotopes
  gecko isotopes linked to body sizes	pre-2010	unclear what years these geckos      came from. Data from Briggs et al. 2012 Plos paper
  
  File: 2012_Palmyra_Plant_Isotopes
  plant isotopes	2012	per island	this is duplicate (with maybe extra) of         2010-2012 plant datasheet
  
  File: 2013-2015 Palmyra_Geograpsus_crab_weights_and_isotopes_data
  Geograpsus isotope and body size data	2013-2015	individuals**
  
  File: 2010_Palmyra_Day_Spider_isotopes
  Neoscona theisi isotope data	2010	pooled
  
  File: 2012_Palmyra_Day_Spider_Isotopes
  2012	Day Spider isotopes	pooled
  
  File: 2010_Palmyra_Insect_Isotopes
  insect isotope data	2010	pooled	crickets in this year were either Cycloptilum    or Ornebius collected during sweeps of understory vegetation
  
  File: 2011_Palmyra_Soil_Isotopes
  soil isotopes	2011	pooled from islet	

  File: 2010_Palmyra_Gecko_Isotopes
  gecko isotopes	2010	both pooled and individual	unclear if there were more this     year? This is a small sample…
			not sure if these were by species or not, seems that they were not                 distinguished this way, so genus is best level of resolution

  File: 2010_Palmyra_Plant_Isotopes
  plant isotope data	2010	this is duplicate data (with maybe more) from the        compiled 2010-2012 data

  File: 2011_Palmyra_Plant_isotopes
  plant isotopes	2011	one plant on a sheet of soil samples


  File: 2012_Palymra_Cane Spider_Isotopes
  2012	Cane spider isotopes	pooled and some individual	pooled isotope data from     2-15 individuals each, with individaul isotope data for 12 Eastern and 10 Kaula     individuals
  
  File: 2011_Palmyra_Cane_Spider_Isotopes
  2011	cane spider isotopes	Individual	only one data point on a sheet with all    soil samples, no islet name

  File: 2011_Palmyra_Gecko_Isotopes
  gecko isotopes	2011	individuals	all run by Nicole Sarto except Sand A-K, which     were run on a different run by AMtK
	unclear if one species or another, genus is best resolution

  File: 2015_Palmyra_Cane_Spider_Isotopes_Body_Size
  cane spider isotopes and body size	2015	individuals	body size is body length     from front of cephalothorax to back of thorax
  
  d15NNorm**	and Wt.% N are Nitrogen vs Air
  d13CNorm**	and Wt.% C are Carbon vs PDB
  

### Other Changes of Note
  Manually combined duplicate values for 2013-2015 Geograpsus crab weight and        isotope sheets and carried over non-duplicate columns to consolidate and preserve     data
  Due to different sample ID conventions, not all sheets in the original file could     be combined
  Combined 2013 values from sheets 'isotopes 2013' and '2013-14' into new file -      2013_Palmyra_Geograpsus_weight_isotopes
  Combined 2014 values from sheets '2013-14' and 'isotopes 2014-2014' into new file - 2014_Palmyra_Geograpsus_weight_Isotopes
  Combined 2015 values from sheets 'isotopes 2014-2015' and '2015' into new file - 2015_Palmyra_Geograpsus_length_isotopes
  Sheets 'weights 2013-14' and 'lengths 2015' could not be consolidated due to sample ID ambiguity


## Set up and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


# Installs and loads packages if the user doesn't have them already installed
if (!require("googledrive")) {install.packages("googledrive"); require("googledrive")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("janitor")) {install.packages("janitor"); require("janitor")}
if (!require("lubridate")) {install.packages("lubridate"); require("lubridate")}

# Optional package loading if you don't want to use the above ^
library(googledrive)
library(here)
library(data.table)
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)


## Sets R not to use scientific notations
options(scipen=999) 

# Create raw data folders - DO I NEED A IF STATEMENT???

if (!dir.exists(here::here("/raw_data"))){
  dir.create(paste0(getwd(),"/raw_data"))
}

## If necesary create a sub directory inside the raw_data. This will be the case for projects that will end up in more than one package.
dir.create(paste0(getwd(),"/raw_data/isotope"))
```

## Googledrive package sign-in

```{r}
# To change accounts on the googledrive package, uninstall then reinstall the package. Run a googledrive function to prompt the sign-in selection screen
remove.packages("googledrive")
install.packages("googledrive")

drive_find(n_max = 10)
```

With set-up complete, now begin opening and cleaning data


## Download data files into local computer

In this case, the project has multiple types of data. Each type of data will be cleaned in its own Rmd and will become its own package. 
Make sure the file path on the code is updated to the correct folder. In this case we are working with the weather data, so data is saved inside the `isotope` folder. Update folder name if you are downloading other type of data.

```{r download files}
# url of folder where the data lives
# Copy paste the url of the folder where the data lives

folder_url <- "https://drive.google.com/drive/u/1/folders/1mWTvcrU-P3tu5brHKXArBbgoV-NbnVzd"

# list of files inside the folder
files <- drive_ls(as_id(folder_url))

## Downloads each file to local computer
purrr::walk2(
    map(files$id, as_id),
    paste0("raw_data/isotope/", files$name),
    drive_download,
    overwrite = TRUE
  )

# Check all files where downloaded
# Count files inside the raw_data folder to make sure the number of files downloaded is what is expected.
raw_data_path <- paste0(getwd(), "/raw_data/isotope")

length(list.files(raw_data_path))

```


## Convert files into a tibble (without renaming) - Not Tested

```{r}

## List wit all csv files in raw data
all_csv <- tibble(list.files(raw_data_path, pattern = ".csv")) %>% 
  rename(file_name = 1) %>% 
  mutate(path = paste0(raw_data_path, "/", file_name),
         n = 1:n(),
         type = "isotope") %>% 
  unite(obj_name, type, n, sep = "_", remove = FALSE)

## function to read csv and clean names as we read the file
read_csv_clean <- function(dataset){
  
  read_csv(dataset) %>% 
    clean_names()
```

Instead of immediately renaming the files, this data package will likely end up being sorted into three separate files - Soil Isotopes, Plant Isotopes, and Animal Isotopes.
Sheets will be cleaned and examined for merging before being renamed to make comparison with the excel files easier.


Next Steps:
Test read code to open the files
Group files based on column similarities and identify which can be merged first
Clean other sheets to allow merging

makning changes to test git
