---
title: "Cleaning Body Size data"
author: "Camila Vargas"
date: "8/12/2021"
output: html_document
---

This script cleans data on Body sizes for taxa including cane spiders, geckos, and other spiders and insects provided by Ana Miller-ter Kuile in July 2021.
Find Miller-ter Kuile raw data [here](https://github.com/anamtk/palmyra/tree/main/raw_data/body_sizes)

## Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(googledrive)
library(here)
library(data.table)
library(tidyverse)
library(readxl)
library(janitor)
library(lubridate)

## Sets R not to use scientific notations
options(scipen=999) 

# Create raw data folders - IS THE IF STATEMENT NECESARY???

if (!dir.exists(here::here("/raw_data"))){
  dir.create(paste0(getwd(),"/raw_data"))
}

## If necesary create a sub directory inside the raw_data. This will be the case for porject that will end up in more than one package.
dir.create(paste0(getwd(),"/raw_data/body_size"))

```



## Download data files into local computer

In this case, the project has multiple types of data. Each type of data will be cleaned in its own Rmd and will become its own package. 
Make sure the file path on the code is updated to the correct folder. In this case we are working with the weather data, so data is saved inside the `weather` folder. Update folder name if you are downloading other type of data.

```{r download files}
# url of folder where the data lives
# Copy pasete the url of the folder where the data lives

folder_url <- "https://drive.google.com/drive/u/1/folders/1Ypg74d5XzyCLG03qEVR1eaJOrvREyRHC"

# list of files inside the folder
files <- drive_ls(as_id(folder_url))

## Download each file to local computer
purrr::walk2(
    map(files$id, as_id),
    paste0("raw_data/body_size/", files$name),
    drive_download,
    overwrite = TRUE)

```

## Check all files were downloaded 
```{r}
## File path to raw data used in this script
raw_data_path <- paste0(getwd(), "/raw_data/body_size")

# Count files inside the raw_data forder to make sure the number of files downloaded is what is expected.
length(list.files(raw_data_path))

```



## Functions and tables with files info
Creating a table with all of our files by type to use as an index to access each file.

```{r read data}

## List wit all csv files in raw data
all_csv <- tibble(list.files(raw_data_path, pattern = ".csv")) %>% 
  rename(file_name = 1) %>% 
  mutate(path = paste0(raw_data_path, "/", file_name),
         n = 1:n(),
         type = "body_size") %>% 
  unite(obj_name, type, n, sep = "_", remove = FALSE)

## List with all xls files
all_xls <- tibble(list.files(raw_data_path, pattern = c("xls", "xlsx"))) %>% 
  rename(file_name = 1) %>% 
  mutate(path = paste0(raw_data_path, "/", file_name),
         n = 1:n(),
         type = "body_size") %>% 
  unite(obj_name, type, n, sep = "_", remove = FALSE)


## function to read csv and clean names as we read the file
read_csv_clean <- function(dataset){
  
  read_csv(dataset) %>% 
    clean_names()
}


```

## Read csv files

The only csv file is Spider Insect mass length

```{r}
## Read in all csv
for (i in all_csv$n){
  
  assign(all_csv$obj_name[i], read_csv_clean(all_csv$path[i]))
}

```

body_size_1 - 2010-2012_Spider_Insect_mass_length - This data set is for the most part organized. It needs some standarization of dates and names.

## Read excel files

**Cane Spider**
```{r}
## chseck sheet names
excel_sheets(all_xls$path[1])

cane_spider_09 <- read_excel(all_xls$path[1], sheet = "Sheet1") %>% 
  clean_names()

## Is this table a summary table from data above - We will not include processed or summary data.
cane_spider_09_2 <- read_excel(all_xls$path[1], sheet = "Sheet2") %>% 
  clean_names()

cane_spider_10 <- read_excel(all_xls$path[2], sheet = "Sheet1") %>% 
  clean_names()


excel_sheets(all_xls$path[6])
cane_spider_15_loc <- read_excel(all_xls$path[6], sheet = "Spider Vial Locations") %>% 
  clean_names()

cane_spider_15_size <- read_excel(all_xls$path[6], sheet = "Spider Sizes") %>% 
  clean_names()

cane_spider_15_me <- read_excel(all_xls$path[6], sheet = "Measurement Errors") %>% 
  clean_names()


```

## Cleaning cane spider data 
```{r}
cs_09_clean <- cane_spider_09 %>% 
  rename(notes = measurement_complications) %>% 
  slice(1:132) %>% 
  mutate(year = 2009,
         date_collected = NA,
         prosoma = NA,
         species = "cane spider") %>% 
  unite(spider_id, site, spider_number, sep = " ", remove = FALSE) %>% 
  select(year, date_collected, species, site, spider_id, length, weight, prosoma, notes)


cs_10_clean <- cane_spider_10 %>% 
  rename(site = island,
         weight = weight_g,
         length = length_mm) %>% 
  mutate(year = 2010,
         date_collected = NA,
         prosoma = NA,
         species = "cane spider", 
         # All measurments that have a not will be converted to NAs
         weight = as.numeric(weight),
         length = as.numeric(length)) %>% 
  unite(spider_id, site, number, sep = " ", remove = FALSE) %>% 
  select(year, date_collected, species, site, spider_id, length, weight, prosoma, notes)

cs_15_clean <-  cane_spider_15_size %>% 
  rename(weight = weight_g,
         length = length_mm,
         prosoma = prosoma_mm,
         notes = note) %>% 
  mutate(year = 2015,
         species = "cane spider") %>% 
  select(year, date_collected, species, site, spider_id, length, weight, prosoma, notes)
  

## Not sure what measuerment errors are but it lookslike for some spiders id there where more than one measurement and the slightly vary from each other. ASK ANA if this is worth including.
## Not clear what the Vial locations are and if we should include this data.
  

cane_spider_all <- bind_rows(cs_09_clean, cs_10_clean, cs_15_clean)

write_csv(cane_spider_all, path = "clean_data/2009_2010_2015_cane_spider_body_size.csv")

```

Notes for data on 2009:
Juiced = fluids coming out from spider's body, may have decreased weight
Squished = organs/solids coming out of body
Broken = head broken off of abdomen, difficult to measure length
Length measured from white stripe in front of eyes to tip of abdomen


**Gecko**
```{r}
excel_sheets(all_xls$path[3])
gecko_10 <- read_excel(all_xls$path[3], sheet = "Sheet1") %>% 
  clean_names()

excel_sheets(all_xls$path[4])
gecko_11 <- read_excel(all_xls$path[4], sheet = "geckos", skip = 1) %>% 
  clean_names()

excel_sheets(all_xls$path[4])
gecko_11_summary <- read_excel(all_xls$path[4], sheet = "survey summary", skip = 1) %>% 
  clean_names()

excel_sheets(all_xls$path[5])
gecko_14 <- read_excel(all_xls$path[5], sheet = "geckos", skip = 1) %>% 
  clean_names()

gecko_14_summary <- read_excel(all_xls$path[5], sheet = "survey summary", skip = 1) %>% 
  clean_names()

excel_sheets(all_xls$path[7])
gecko_15_surv <- read_excel(all_xls$path[7], sheet = "Survey Data") %>% 
  clean_names()

gecko_15_me <- read_excel(all_xls$path[7], sheet = "Gecko Measurements") %>% 
  clean_names()


```


## Clening Gecko data

Data for years: 2010, 2011, 2014, and 2015

We have length and weight meassurments and also suvey details (_summary or _surv)


```{r}
g_10_clean <- gecko_10 %>% 
  rename(length = body_length) %>% 
  mutate(year = 2010,
         site = NA,
         date = NA,
         species = "gecko",
         notes = "From an email by Alex Mclnturff")


g_11_clean <- gecko_11 %>% 
  rename(site = island,
         length = tot_length,
         weight = weight_g) %>% 
  mutate(year = 2011) %>% 
  ## all "NA" will become real NAs
  mutate_at(vars(length, weight, eggs, svl, tail_collected), as.numeric) %>% 
  unite(gecko_id, site, id_number, sep = " ", remove = FALSE) %>% 
  select(year, date, species, site, gecko_id, length, weight, svl,eggs, sex, tail_collected, notes)

## Notes: some tail samples were too small for isotope samples, so those islands were revisited.
## During those initial surveys, SVL was not measured, and sex and reproductive status of individuals was not thoroughly checked, so those cells where that info is uncertain are marked with "NA".
##  At some islands, extra geckos (past number needed for isotopes) were collected during the survey period.  To save time, the geckos were subsampled, and only 12-19 were fully measured and tails collected, though all geckos collected were id'd to species


g_14_clean <- gecko_14 %>% 
  rename(site = island,
         weight = weight_g) %>%
  mutate(year = 2014,
         length = NA) %>% 
  mutate_at(vars(length, weight, svl), as.numeric) %>% 
  unite(gecko_id, site, id_number, sep = " ", remove = FALSE) %>%
  select(year, date, species, site, gecko_id, length, weight, svl,eggs, sex, notes)
  

g_15_clean <- gecko_15_me %>%
  separate(col = gecko_id,
           into = c("site", "id"),
           sep = " ",
           remove = FALSE) %>% 
  mutate(year = 2015,
         lenght = NA,
         date = case_when(site %in% c("Strawn", "Aviation") ~ "2015-07-11",
                          site %in% c("Frigate", "Portsmouth", "Eastern") ~ "2015-07-12",
                          T ~ "2015-07-13") %>% as.Date(.))%>% 
  rename(weight = wt_g, 
         svl = svl_mm) %>% 
  select(year, date, species, site, gecko_id, weight, svl, notes)


## Combine all years of Geko data

gecko_all <- bind_rows(g_10_clean, g_11_clean, g_14_clean, g_15_clean) %>% 
  mutate(species_note = case_when(str_detect(species, "\\?") ~ "not clear"),
         species = str_remove_all(species, "\\?") %>% trimws(.),
         species_note = ifelse(species == "" & species_note == "not clear", "unknown", species_note),
         species = ifelse(species %in% c("hybrid", "Hybird", "Hybrid"), "hybrid", species),
         species_note = case_when(species == "unknown, probably L. lug" ~ "not clear",
                                  species == "male of new species (or hybrid, photo taken)" ~ "unknown",
                                  species == "unknown - either dark morph L. lug or house gecko" ~ "unknown - either dark morph L. lug or house gecko",
                                  T ~ species_note),
         species = case_when(species == "L." ~ "Lug",
                             species %in% c("L.lug", "unknown, probably L. lug") ~ "L. lug",
                             species %in% c("male of new species (or hybrid, photo taken)", "unknown - either dark morph L. lug or house gecko") ~ "",
                             T ~ species),
         species = ifelse(species %in% c("NA", ""), NA, species),
         site = case_when(site == "Whip" ~ "Whippoorwill",
                          T ~ site),
         sex = case_when(sex %in% c("F", "M", NA) ~ sex,
                         sex == "male" ~ "M")) %>%
  select(year, date, species, species_note, site, gecko_id, length, weight, svl, eggs, sex, tail_collected, notes)
                             

write_csv(gecko_all, "clean_data/2010_2015_gecko_body_size.csv")

```

## Cleaning insect data

```{r}
insect_clean <- body_size_1 %>% 
  mutate(year = case_when(date == "2012" ~ 2012,
                          T ~ 2010),
         date = ifelse(date == "2012", NA, date) %>% as.Date(., format = "%m/%d/%y"),
         number = case_when(number == "252/282" ~ "252",
                            T ~ number) %>% as.numeric(.),
         island = case_when(island == "NF" ~ "North Fighter",
                            island == "SF" ~ "South Fighter",
                            island == "P.mouth" ~ "Portsmouth",
                            island == "WHIP" ~ "Whippoorwill",
                            T ~ island)) %>% 
  rename(site = island,
         weight = weight_mg,
         length = length_mm) %>% 
  select(year, date, everyhing())



```
